\Title

### 2023-07-12 Grimaldi-etal-BiolCybernetics.tex
methods

The precise timing of spikes emitted by neurons plays a crucial role in shaping the response of efferent biological neurons. This temporal aspect of neural activity holds significant importance in understanding information processing in neurobiology, especially concerning the output of neuromorphic hardware like event-based cameras. However, many artificial neural models disregard this critical temporal dimension of neural activity. In this study, we present a model designed to efficiently detect temporal spiking motifs using a layer of spiking neurons equipped with heterogeneous synaptic delays. Our model capitalizes on the diverse synaptic delays present on the dendritic tree, enabling specific arrangements of temporally precise synaptic inputs to synchronize upon reaching the basal dendritic tree. We formalize this process as a time-invariant logistic regression that can be trained using labeled data. To demonstrate its practical efficiency, we apply the model to naturalistic videos transformed into event streams, resembling the output of the biological retina or event-based cameras. To assess the robustness of the model in detecting visual motion, we conduct experiments by selectively pruning weights and demonstrate that the model remains efficient even under significantly reduced workloads. In conclusion, by offering a comprehensive, event-driven computational building block, heterogeneous delays has the potential to enhance the performance of future spiking neural network algorithms, particularly within the realm of neuromorphic chips.

@@

In this paper, we test whether the HD-SNN model is able to efficiently learn to solve a motion detection task defined by a realistic event-based data stream. This type of signal is commonly captured by an event-based camera, or Dynamic Vision Sensor (DVS), and is inspired by the signal sent from the retinal ganglion cells through the optic nerve to the visual cortex. The events are binary in nature and should be sufficient to perform the motion detection task in a fast and efficient manner. Here, we first describe the task definition, then the HD-SNN model we use for motion inference, and finally how we train it.

@@

        In response to an instance of the event-based input (Left), we show the neural activity of the HD-SNN model (Right). For visualization purposes, we show only one spatio-temporal slice for a given vertical position ($y = 32$), for a quarter of horizontal positions. The input spiking activity consists of ON and OFF spikes (as in~\seeFig{motion_task}) and illustrates the switching of the naturalistic event-based stream from one motion to another, as denoted by the colored dots above the graph (with the class of each output neuron is represented by a similar color as in \seeFig{motion_task}). The output activity consists of an analog component corresponding to the evidence accumulated by the model on the spatio-temporal kernels, and a spiking part (overlaid on this activity as vertical bars) which corresponds to instants when this evidence is above the spiking threshold. Note that this activity is consistent with the motion of the input stream. Finally, an accuracy can be computed by comparing the ground truth motion in the input movie with that predicted by the model.

@@

After training our MLR model, we obtain spatiotemporal kernels corresponding to the weights associated with the heterogeneous delays of our layer of spiking neurons, which can be used for detection. For this, we quantify its ability to categorize different motions, i.e. on event streams for which the ground truth motion is known at each instant. When applied to new instances of the input movies, the model develops a neural activity which may be used to infer the correct motion~\seeFig{activity} and from which we may deduce an accuracy value. This accuracy was computed on a novel dataset of $200$ novel movies. The accuracy computed on the test set was approximately $91\%$ (with a chance level of $1 / 12/ 3 \approx 2\%$). %Moreover, we observed that the classification performance obtained for the model using time surfaces as in~\citep{grimaldi_robust_2022} gives similar results as our method.

>>>>

The precise timing of spikes emitted by neurons plays a crucial role in shaping the response of efferent biological neurons. This temporal dimension of neural activity holds significant importance in understanding information processing in neurobiology, especially concerning the output of neuromorphic hardware, such as event-based cameras. Nonetheless, many artificial neural models disregard this critical temporal dimension of neural activity. In this study, we present a model designed to efficiently detect temporal spiking motifs using a layer of spiking neurons equipped with heterogeneous synaptic delays. Our model capitalizes on the diverse synaptic delays present on the dendritic tree, enabling specific arrangements of temporally precise synaptic inputs to synchronize upon reaching the basal dendritic tree. We formalize this process as a time-invariant logistic regression, which can be trained using labeled data. To demonstrate its practical efficacy, we apply the model to naturalistic videos transformed into event streams, simulating the output of the biological retina or event-based cameras. To evaluate the robustness of the model in detecting visual motion, we conduct experiments by selectively pruning weights and demonstrate that the model remains efficient even under significantly reduced workloads. In conclusion, by providing a comprehensive, event-driven computational building block, the incorporation of heterogeneous delays has the potential to enhance the performance of future spiking neural network algorithms, particularly in the context of neuromorphic chips.

@@

This paper aims to investigate the capability of the HD-SNN model to effectively learn and solve a motion detection task using realistic event-based data streams, as typically captured by event-based cameras, also known as Dynamic Vision Sensors (DVS). DVS are designed to mimic the signal transmitted from retinal ganglion cells to the visual cortex through the optic nerve. The events in these data streams are binary in nature and should provide sufficient information for performing fast and efficient motion detection. In this study, we first outline the task definition, specifying the requirements for motion detection. Subsequently, we describe the HD-SNN model employed for inferring motion, highlighting its key characteristics and architecture. Lastly, we elaborate on the training procedure employed to train the HD-SNN model for the specific motion detection task.

@@

        In response to a specific event-based input instance (Left), we present the neural activity of the HD-SNN model (Right). To aid visualization, we display a single spatio-temporal slice for a given vertical position ($y = 32$) and $10$ horizontal positions. The input spiking activity comprises ON and OFF spikes, as explained in Fig.~\ref{fig:motion_task}, showcasing the switching within the naturalistic event-based stream from one motion to another due to saccades. The dots above the graph indicate the corresponding motion class at each instant (the motion being represented by the matching color). The output activity consists of two components. Firstly, there is an analog component that corresponds to the evidence accumulated by the model on the spatio-temporal kernels. Secondly, there is a spiking component represented by vertical bars superimposed on the analog activity. These bars signify moments when the evidence surpasses the spiking threshold. Importantly, this activity aligns with the motion depicted in the input stream. Finally, it is possible to compute the accuracy by comparing the ground truth motion in the input video with the motion predicted by the model (as represented by the colored dots).

@@

After training our MLR model, we obtain spatiotemporal kernels corresponding to the weights associated with the heterogeneous delays of our layer of spiking neurons, which can be used for detection. For this, we quantify its ability to categorize different motions, i.e. on event streams for which the ground truth motion is known at each instant. When applied to new instances of the input movies, the model develops a neural activity which may be used to infer the correct motion~(\seeFig{activity}) and from which we may deduce an accuracy value. This accuracy was computed on a novel dataset of $200$ novel movies. The accuracy computed on the test set was approximately $91\%$ (with a chance level of $1 / 12/ 3 \approx 2\%$). %Moreover, we observed that the classification performance obtained for the model using time surfaces as in~\citep{grimaldi_robust_2022} gives similar results as our method.
